CF-SuGAR Power Calculations

### 2025-03-19 Christine Chan/CF-SuGAR/power_calculations.qmd
power calculations to do list

```{r}
# Origin for survival analyses
og <- ymd("2003-01-01")
# Need to use a dataframe not a tibble
annual <- as.data.frame(annual)
annual$tstart <- as.numeric(annual$tstart) - as.numeric(og)
annual$tstop <- as.numeric(annual$tstop) - as.numeric(og)
# Formula for base model
xvars <- c(
  "age_group", "Gender", "Race", "Hispanicrace", "bmi", "MutClass",
  "A_IsOnEnzymes", "tube_feeding", "liver_disease", "A_FEV1_pct_predicted",
  "NumPulmExacerbation", "A_corticosteroids1", "A_hgba1c"
)
y <- "Surv(time = tstart, time2 = tstop, event = as.numeric(CFRD) == 2)"
f <- as.formula(paste0(y, "~", paste0(xvars, collapse = "+")))
cox_res <- coxph(f, data = annual, cluster = eDWID)
cox.zph(cox_res)
# Forest plot and coefficients table
cox_res %>%
  tbl_regression(exponentiate = T)
ggcoef_table(cox_res,
  show_p_values = F, add_reference_rows = T, exponentiate = T,
  signif_stars = F, colour = F
)
```

@@

```{r}
# Formula for base model
xvars <- c(
  "age_group", "Gender", "bmi", "MutClass", "A_FEV1_pct_predicted", "A_hgba1c"
)
y <- "Surv(time = tstart, time2 = tstop, event = as.numeric(CFRD) == 2)"
f <- as.formula(paste0(y, "~", paste0(xvars, collapse = "+")))
cox_res <- coxph(f, data = annual, cluster = eDWID)
cox.zph(cox_res)
# Forest plot and coefficients table
cox_res %>%
  tbl_regression(exponentiate = T)
ggcoef_table(cox_res,
  show_p_values = F, add_reference_rows = T, exponentiate = T,
  signif_stars = F, colour = F
)
```

@@

```{r}
# CPH model for baseline 1 hour glucose
cox1hr <- coxph(
  Surv(Follow_up_time_yrs, CFRD) ~ age + Sex + BMI + class1 + FEV1.per.pred +
    A1C,
  data = baseline_mult
)
# Forest plot and coefficients table
cox1hr %>%
  tbl_regression(exponentiate = T)
ggcoef_table(cox1hr,
  show_p_values = F, add_reference_rows = T, exponentiate = T,
  signif_stars = F, colour = F
)
```

@@

```{r}
# CPH model for baseline 1 hour glucose
cox1hr <- coxph(
  Surv(Follow_up_time_yrs, CFRD) ~ age + Sex + BMI + class1 + FEV1.per.pred +
    A1C + Baseline_1hr_glucose,
  data = baseline_mult
)
# Forest plot and coefficients table
cox1hr %>%
  tbl_regression(exponentiate = T)
ggcoef_table(cox1hr,
  show_p_values = F, add_reference_rows = T, exponentiate = T,
  signif_stars = F, colour = F
)
```

@@

>>>>

```{r}
# Origin for survival analyses
og <- ymd("2003-01-01")
# Need to use a dataframe not a tibble
annual <- as.data.frame(annual)
annual$tstart <- as.numeric(annual$tstart) - as.numeric(og)
annual$tstop <- as.numeric(annual$tstop) - as.numeric(og)
# Formula for base model
xvars <- c(
  "age_group", "Gender", "Race", "Hispanicrace", "bmi", "MutClass",
  "A_IsOnEnzymes", "tube_feeding", "liver_disease", "A_FEV1_pct_predicted",
  "NumPulmExacerbation", "A_corticosteroids1"
)
y <- "Surv(time = tstart, time2 = tstop, event = as.numeric(CFRD) == 2)"
f <- as.formula(paste0(y, "~", paste0(xvars, collapse = "+")))
cox_res <- coxph(f, data = annual, cluster = eDWID)
cox.zph(cox_res)
# Forest plot and coefficients table
cox_res %>%
  tbl_regression(exponentiate = T)
ggcoef_table(cox_res,
  show_p_values = F, add_reference_rows = T, exponentiate = T,
  signif_stars = F, colour = F
)
```

@@

```{r}
annual = annual%>% filter(Age_YrEnd<=22)
annual$age_group = cut(annual$Age_YrEnd,c(7,11,22))
# Formula for base model
xvars <- c(
  "age_group", "Gender"
)
y <- "Surv(time = tstart, time2 = tstop, event = as.numeric(CFRD) == 2)"
f <- as.formula(paste0(y, "~", paste0(xvars, collapse = "+")))
cox_res <- coxph(f, data = annual, cluster = eDWID)
cox.zph(cox_res)
# Forest plot and coefficients table
cox_res %>%
  tbl_regression(exponentiate = T)
ggcoef_table(cox_res,
  show_p_values = F, add_reference_rows = T, exponentiate = T,
  signif_stars = F, colour = F
)
```

@@

```{r}
# CPH model for baseline 1 hour glucose
cox1hr <- coxph(
  Surv(Follow_up_time_yrs, CFRD) ~ age + Sex ,
  data = baseline_mult
)
# Forest plot and coefficients table
cox1hr %>%
  tbl_regression(exponentiate = T)
ggcoef_table(cox1hr,
  show_p_values = F, add_reference_rows = T, exponentiate = T,
  signif_stars = F, colour = F
)
```

@@

```{r}
# CPH model for baseline 1 hour glucose
baseline_mult$age_group = cut(baseline_mult$age,c(7,11,22))

cox1hr <- coxph(
  Surv(Follow_up_time_yrs, CFRD) ~ age*Sex,
  data = baseline_mult
)

# Forest plot and coefficients table
cox1hr %>%
  tbl_regression(exponentiate = T)
ggcoef_table(cox1hr,
  show_p_values = F, add_reference_rows = T, exponentiate = T,
  signif_stars = F, colour = F
)
```

@@

To Do

- Plot HbA1c by age for those who do develop CFRD
- Take out HbA1c for base model
- Include generalized R2 for the models
- Split age into equal tertiles
  2 x 3 table with rows for diabetes status and age as columns, cells are numbers

- PROC power has a cox reg option
  - Base model is from the registry 
  - 0.5, 1, 2 times the 1 hr glucose point estimate from Lorenz et al.
  - 80, 85, 90% power

- Mixed model power calculations for aim 3
  - Repeated measures with FEV1 and BMI
  - Mutation class and sex as predictors
